\LARGE \bf Reactive Trajectory Planning and Trajectory Tracking for Pedestrian-Aware Autonomous Driving in Urban Environments

### 2016-01-23 root.tex
removed dangling reference to deleted figure

Our approach to fitting appropriate jerk profiles to complex sub-paths is to further break each subpath into segments which have very simple shape.
Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
Each segment will be defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ for must be defined for each segment.

@@

% How we actually did this.
For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$, where $N_p$ is the number of waypoints in the subpath.
The value $\kappa_k$ denotes the path curvature at each point.
For each \textbf{??? sub-path ??} point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.

@@

The problem is defined by the vector of known values $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
Final forward acceleration is set to zero.
If $\mathbf{a}_m = [a^+_m , a^-_m]$ and $\mathbf{j}_m = [j^+_m , j^-_m]$. are known, then a profile with $M$ jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
Jerk and peak acceleration may be different when reducing speed versus increasing speed to more closely mirror human driving behavior.
As we show later, peak acceleration and jerk must sometimes be modified to obtain physically possible time intervals.
To enable this, ranges of acceptable longitudinal jerk and acceleration $\mathbf{a}^{max}_m = [a^{+,max}_m , a^{-,max}_m]$ and $\mathbf{j}^{max}_m = [j^{+,max}_m , j^{-,max}_m]$ are set according to the relation in Eqs. \ref{eq:am} and \ref{eq:jm}.
This is computed for every point along the subpath.
In the simplest case, one value for each of the quantities expressed in these relations may be chosen for the entire path.

@@

There are many cases in which this solution will fail.
For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
We propose to define 4 other basic profiles, and our algorithm judiciously chooses the appropriate profile for every segment defined by a vector $\mathbf{b}$.

\textbf{RG: does not belong here: The other profiles are explained below in Section \ref{sec:jerkprofiles}.
A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k, k = 1 ... N_s$ and can be solved individually in sequence. Determination of where to break a path into segments is discussed below in Section \ref{sec:pathsegmentation}. }

There are a total of five potential jerk profiles which have been created to cover all feasible values in $\mathbf{b}$.
All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
To find closed-form solutions for each, a symbolic solver is used.
The target values for each profile are the time steps between jerk changes.
These are solved by plugging in $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$.

For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
This results in a negative time interval for that phase.
In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
Specifically, jerk may be increased while peak acceleration is decreased.
This process is referred to herein as ``accel-jerk tuning''
If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.

To decide which profile to employ, a cascade approach is used.
Analytic values of time intervals are evaluated for each profile by plugging in values of $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$ until all time intervals are positive and real.
The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.

@@

A change from one speed to another.
All of the above profiles are composed of some combination of a 3-phase profile and a $j_k = 0$ phase.
This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial speed, so fitting one of the two 4-phase profiles has been attempted and subsequently failed.

\subsection{Stitching segment profiles together} \label{sec:unifiedsolution}

At this point, each sub-path has been divided into segments and conditions have been computed such that an analytically derived plan can be fit to each segment individually.
Now a reference trajectory must be generated for the entire sub-path.
Continuity of position, speed, and acceleration is the primary goal of this process.

A jerk vs. time profile is fit to each segment using the algorithm in Section \ref{sec:jerkprofiles}.
Speed and acceleration must be continuous between each segment, but only the corresponding speed ceilings have been found, so the start and end conditions for each segment must be determined.
The values of $a_i$ and $v_i$ for the first segment is set to be the value which is currently reported from sensor data.
Since the final acceleration is set to zero for all profiles, each subsequent segment may begin with $a_i = 0$.
For the final segment, setting $v_f = 0$ is prudent unless the scenario demands otherwise.
Algorithm~\ref{alg:segmentspeedboundaryconditions} is used to set boundary speed conditions for the interior segments.

\begin{figure}
  \begin{algorithmic}[1]
    \Procedure{SegmentBoundarySpeeds}{$v_m$}
      \State $\mathbf{v}_p \gets \mathbf{v}_m$ \Comment{Initialize targets as ceilings}
      \State $\mathbf{v}_{p,0} \gets v_i$ \Comment{Set starting speed to current}
      \State $\mathbf{v}_p = [\mathbf{v}_p, 0]$ \Comment{Add final speed of 0}
      \For{$i = 1, ..., N_s-1$}
        \If{$v_{m,i+1} > v_{m,i}$}
          \State $v_{p,i+1} \gets v_{m,i+1}$
        \Else
          \State $v_{p,i+1} \gets v_{m,i}$
        \EndIf
      \EndFor
    \EndProcedure
  \end{algorithmic}
  \caption{
    Algorithm to set speed boundary conditions for interior path segments.
    Having consolidated the waypoints in to a set of path segments, the maximum allowable speed for each segment, $v_{m,i}$ is known.
    This algorithm uses the set of $\mathbf{v_{m}}$ for all segments to choose the highest allowable start and end speeds for each path segment.
  }
\label{alg:segmentspeedboundaryconditions}
\end{figure}

% talk about control points
Now each segment is solved individually.
In the case where a segment requires a speed change which is infeasible given limits on acceleration and jerk, then the final speed for that segment is planned to be the closes possible value.
When these discrepancies in the feasible $v_f$ for a segment arise, the corresponding value in $v_p$ is updated so that the new conditions may be accounted for in the subsequent segment.

At the end of the process, there will be a set of $N_{cp}$ ``control points''.
A control point is a vector $[t_{ref}, s, v_{B,x}, a_{B,x}, j_{B,x}]$ which describes the longitudinal kinematic state of the target vehicle at the moment of an infinite jounce jerk change.
Fitting a jerk profile to each segment results in control points which must be stitched into the control points for the preceding segments.
While no further information than initial conditions, time durations, and jerk values is needed (simple integration would yield lower derivatives of position), this information is already available since it must be solved for each segment.
This extra information may be used as a redundancy layer to verify that no errors occurred.
For each segment with $M$ jerk phases, the following is produced:
\begin{itemize}
  \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
  \item $M+1$ values of distance elapsed within the segment $\Delta s$. These must be shifted forward by the last $s$ value from the previous segment to translate them into the 1D path domain. Additionally, the first value must be erased.
  \item $M+1$ values of $v_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
  \item $M+1$ values of $a_{B,x}$. The first value is checked against the last value from the previous segment to verify equality, then it is erased. These need no translation.
  \item $M$ jerk values, which may be appended directly to the preceding jerk values
\end{itemize}

\textbf{RG: please explain reference time, $s$ and $\Delta s$ ????????}

After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
Since $j(t)$ is piecewise constant, it may be integrated to determine an instantaneous trajectory at any moment in time.
To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.

@@

After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
This is depicted in Fig. \ref{fig:pathtostop}
The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.

@@

Now the reference time corresponding to the vehicle's location $t_{veh}$ is needed to look up $v_{B,x}(t_{veh})$, $a_{B,x}(t_{veh})$, and $j_{B,x}(t_{veh})$.
Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$, F may be approximated \textbf {?????? as cubic if a spline is ???? } being constructed from knots which have a very high resolution.
In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:unifiedsolution}, and it is now resampled at $\Delta t = 0.01 sec$.
This yields two vectors $s'$ and $t'$ which are then used to create a univariate cubic $t(s)$ spline.
Evaluating the $t(s_{veh})$ then yields the current reference time.

>>>>

Our approach to fitting appropriate jerk profiles to complex sub-paths is to further break each subpath into segments which have very simple shape.
Once this is done, each segment can be compared to a library of possible jerk profiles to find a good match.
Each segment will be defined by the vector $\mathbf{b}  = [v_i, a_i, L, v_m, v_f]$, being initial speed, initial acceleration, the total distance, speed limit, and the final speed, respectively.
Additionally, the desired positive and negative acceleration $\mathbf{a}_m = [a^+_m , a^-_m]$ and the desired positive and negative jerk $\mathbf{j}_m = [j^+_m , j^-_m]$ for must be defined for each segment.
A path is then a set of $N_s$ segments, each of which is defined by a set of parameters $\mathbf{b}_k$, $\mathbf{a}_{m,k}$, and $\mathbf{j}_{m,k}$ where $k = 1 ... N_s$ and can be solved individually in sequence.

@@

% How we actually did this.
For the purposes of developing a simple example application, two constraints will be used for $v_{m}$: legal speed limit $SL$ and lateral acceleration limit $a_{B,y}^{max}$. 
Lateral acceleration constraints are translated into longitudinal speed constraints using the approximation $v_{m,k}(a_{B,y}) = \sqrt{a_{B,y}^{max}/\kappa_k}$ for $k = 1, ..., N_p$, where $N_p$ is the number of waypoints in the subpath.
The value $\kappa_k$ denotes the path curvature at each point.
For each sub-path point, the value $v_{m,k}(a_{B,y}^{max})$ denotes the speed necessary to obtain the limit lateral acceleration, denoted by $a^{max}_{B,y}$.

@@

Final forward acceleration is set to zero.
A profile with $M$ constant jerk phases (7 in this case) can be generated by solving for the time intervals $[\Delta t_1, ..., \Delta t_M]$.
As we show later, $\mathbf{a}_m$, and $\mathbf{j}_m$ must sometimes be modified to obtain physically possible time intervals.
To enable this, ranges of acceptable longitudinal jerk and acceleration $\mathbf{a}^{max}_m = [a^{+,max}_m , a^{-,max}_m]$ and $\mathbf{j}^{max}_m = [j^{+,max}_m , j^{-,max}_m]$ are set according to the relation in Eqs. \ref{eq:am} and \ref{eq:jm}.
In the simplest case, one value for each of the quantities expressed in these relations may be chosen for the entire path.

@@

There are many cases in which this solution will fail.
For instance, if $L$ is sufficiently short and $v_m$ is sufficiently high, then it will not be possible to ever reach $v_m$ with realistic values of $\mathbf{a}_m$ and $\mathbf{j}_m$.
We propose to define 4 other basic profiles to cover all feasible values in $\mathbf{b}$., and our algorithm judiciously chooses the appropriate profile.
The other profiles are explained below.

All impossible scenarios are automatically ruled out (e.g., end speed above speed limit, traveling speed of zero, negative speeds in $\mathbf{b}$, etc).
It should be noted that traveling backward (i.e., in reverse gear) is not supported in this work.
To find closed-form solutions for $[\Delta t_1, ..., \Delta t_M]$, given $\mathbf{b}$, $\mathbf{a}_m$, and $\mathbf{j}_m$, a symbolic solver was used.

For each of the profiles, it is possible that the chosen values of $\mathbf{a}_m$ and $\mathbf{j}_m$ will not be feasible for a desired speed at the end of the phase.
This results in a negative time interval for one or more phases.
In many cases, modifying $\mathbf{a}_m$ and $\mathbf{j}_m$ such that they still lie within $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$ will remedy this.
Specifically, jerk may be increased while peak acceleration is decreased.
This process is referred to herein as ``accel-jerk tuning''
If no acceptable values are found within the valid ranges of $\mathbf{a}^{max}_m$ and $\mathbf{j}^{max}_m$, a new jerk profile must be chosen.

To decide which profile to employ, a cascade approach is used.
Analytic values of time intervals are evaluated for each profile by plugging in new values of $\mathbf{a}_m$ and $\mathbf{j}_m$ until all time intervals are positive and real.
The order in which they are attempted is: 7 $\rightarrow$ 6 $\rightarrow$ 4 / 4R $\rightarrow$ 3.
Each is described below.

@@

A change from one speed to another.
All of the above profiles are composed of some combination of a 3 phase profile and a $j = 0$ phase.
This presents a difficulty in that either $L$ or $v_f$ may be enforced, but not both (there is no solution to enforce both).
As such, this profile is chosen only for a segment in which it is impossible to completely arrive at the end speed given the initial conditions.
When these discrepancies in the feasible $v_f$ for a segment arise, the corresponding value in $v_p$ is updated so that the new conditions may be accounted for in the subsequent segment.

% \subsection{Stitching Segment Solutions Together} \label{sec:unifiedsolution}

At this point, the subpath has been divided into segments and an analytically derived plan has been fit to each segment individually.
Now a unified reference trajectory must be constructed by stitching these together.

For each segment with $M$ jerk phases, the following values are computed, which correspond to the instants at which jerk changes:
\begin{itemize}
  \item $M$ time durations $\Delta t_{ref}$. These must be added to the last time value from the previous segment to translate them into reference time.
  \item $M$ jerk values $j(t_{ref})$, which may be appended directly to the preceding jerk values
\end{itemize}
It is now possible to integrate $j(t)$ to determine an instantaneous trajectory at any moment in time.

% After stitching together control point vectors for time, position, and its derivatives, all information required to have a trajectory plan is now in place.
% To aid in this, univariate splines are created with time as the independent variable, using the control points as knots.
% Position along the path as a function of reference plan time, $s(t)$, takes the form of a cubic Hermite spline, constructed using the speed control points $v_{B,x}(t)$ to set the derivatives.
% Speed along the path $v_{B,x}(t)$ takes the form of a quadratic Hermite spline, with the acceleration control points $a_{B,x}(t)$ used to set the derivatives at every knot.
% Acceleration $a_{B,x}(t)$ is piecewise linear, so lookup is a trivial matter.
% The same applies to $j_{B,x}(t)$, which may be implemented with a simple table.

% However, the time at which the trajectory plan must be referenced is not always known.
% The solution to this problem is known as trajectory tracking control, and will be discussed in Sec. \ref{sec:trajectorytracking}.

@@

% After this, the spline creation procedure detailed in Sec. \ref{sec:unifiedsolution} is performed.
The path must now be truncated for 1D $\leftrightarrow$ 2D transformation.
Within the trajectory tracker, all portions of the path leading up to the location of RSTOP triggering are removed, as well as the portions after the location which corresponds to $s_{veh} + \Delta s_{RSTOP}$
% This is depicted in Fig. \ref{fig:pathtostop}
The vehicle now begins sampling from this trajectory and is executing a smooth reactive stop.

@@

Now the reference time corresponding to the vehicle's location $t_{veh}$ is needed to look up $v_{B,x}(t_{veh})$, $a_{B,x}(t_{veh})$, and $j_{B,x}(t_{veh})$.
Additionally, it is also needed to sample $s(t)$ at intervals ahead of the vehicle.
Since it is known that $s(t)$ is piecewise cubic, reference time as a function of vehicle position along the path, $t(s)$,F may be approximated as cubic if a spline is being constructed from knots which have a very high resolution.
In this case, an $s(t)$ spline has already been created using the control points from the unified planning process described in Section~\ref{sec:jerkprofiles}, and it is now resampled at $\Delta t = 0.01 sec$.
This yields two vectors $s'$ and $t'$ which are then used to create a univariate cubic $t(s)$ spline.
Evaluating the $t(s_{veh})$ then yields the current reference time.
